Dependent Object Types

### 2012-08-09 article.tex
fix ref
revised new

% misc identifiers
\newcommand{\dom}{\mbox{\sl dom}}
\newcommand{\fn}{\mbox{\sl fn}}
\newcommand{\bn}{\mbox{\sl bn}}
\newcommand{\sig}{\mbox{\sl sig}}
\newcommand{\IF}{\mbox{\mathem if}}
\newcommand{\OTHERWISE}{\mbox{\mathem otherwise}}
\newcommand{\expand}{\prec}
\newcommand{\weakexpand}{\prec^W}
\newcommand{\spcomma}{~,~}
\newcommand{\sat}{{\bf\;sat\;}}

@@

We propose DOT as a core calculus for path-dependent types. We present
the calculus formally in section \ref{pres} and through examples in
section~\ref{dot-examples}. Though we show that the calculus does not
satisfy the standard theorem of preservation in
section~\ref{dot-preservation}, we contribute a proof sketch of type
safety using logical relations in section~\ref{dot-type-safety}. In
section~\ref{discussion}, we discuss choices and variants of the
calculus, as well as related work, and conclude in
section~\ref{conclusion}.

@@

\boxfig{dot-one}{The DOT Calculus : Syntax, Reduction, Type Assignment}{
{\bf Syntax}\medskip
\begin{center}    
$\ba{l@{\hspace{0.2mm}}|@{\hspace{0.2mm}}l}
\ba[t]{l@{\hspace{10mm}}l}
x, y, z    & \lindent{\mbox{Variable}} \\
l          & \lindent{\mbox{Value label}}\\
m          & \lindent{\mbox{Method label}}\\[0.2em]
v ::=      & \lindent{\mbox{Value}} \\
\gap x     & \mbox{variable} \\[0.2em]
t ::=      & \lindent{\mbox{Term}} \\
\gap v     & \mbox{value} \\
\gap \new x c t & \mbox{new instance} \\
\gap t.l  & \mbox{field selection} \\
\gap \mapp t m t  & \mbox{method invocation} \\[0.2em]
p ::= & \lindent \mbox{Path} \\
\gap x & \mbox{variable} \\
\gap p.l & \mbox{selection} \\
c ::= T_c \ldefs{\seq{d}} & \lindent{\mbox{Constructor}} \\[0.2em]
d ::= & \lindent{\mbox{Initialization}}\\
\gap l = v & \mbox{field initialization}\\
\gap m(x) = t & \mbox{method initialization}\\
s      ::= \seq{x \mapsto c} & \lindent\mbox{Store}
\ea
&
\ba[t]{l@{\hspace{10mm}}l}
L ::=      & \lindent{\mbox{Type label}} \\
\gap L_c   & \mbox{class label} \\
\gap L_a   & \mbox{abstract type label} \\[0.2em]
S,T,U,V,W ::= & \lindent\mbox{Type}\\
\gap p.L & \mbox{type selection} \\
\gap T \refine z {\seq D} & \mbox{refinement} \\
\gap T \tand T & \mbox{intersection type} \\
\gap T \tor T & \mbox{union type} \\
\gap \Top  & \mbox{top type} \\
\gap \Bot  & \mbox{bottom type} \\[0.2em]
S_c, T_c ::= & \lindent \mbox{Concrete type} \\
\multicolumn{2}{l}{\gap p.L_c ~|~ T_c \refine z {\seq D} ~|~ T_c \wedge T_c  ~|~ \Top} \\[0.2em]
D ::= & \lindent\mbox{Declaration} \\
\gap \Ldecl L S U & \mbox{type declaration} \\
\gap \ldecl l T   & \mbox{value declaration} \\
\gap \mdecl m S U & \mbox{method declaration}\\
\ \\
\Gamma ::= \seq{x \typ T} & \lindent\mbox{Environment}
\ea
\ea$
\end{center}
\medskip

@@

\infrule[\textsc{new}]
{
y \notin \fn(T') \\
\Gamma \ts T_c \wfe \spcomma T_c \expand_y \seq{D} \\
\Gamma \envplus{y: T_c} \ts \seq{D \sat \seq{d}} \spcomma {t'} \typ {T'}}
{\Gamma \ts \new y {T_c \ldefs{\seq{d}}} {t'} \typ T'}

@@

\begin{multicols}{2}[\judgement{Satisfied Declaration}{\fbox{$\Gamma \ts D \sat \seq{d}$}}]

\infrule[\textsc{tdecl-sat}]
{\Gamma \ts S \sub U}
{\Gamma \ts \Ldecl L S U \sat \seq{d}}

\infrule[\textsc{vdecl-sat}]
{\Gamma \ts v \typ {V'} \spcomma {V'} \sub V}
{\Gamma \ts \ldecl l V \sat l = v,\seq{d}}

\infrule[\textsc{mdecl-sat}]
{\Gamma \ts S \wfe\\
 \Gamma \envplus{x : S} \ts t \typ {T'} \spcomma {T'} \sub T}
{\Gamma \ts \mdecl m S T \sat m(x) = t,\seq{d}}

>>>>

% misc identifiers
\newcommand{\dom}{\mbox{\sl dom}}
\newcommand{\fn}{\mbox{\sl fn}}
\newcommand{\bn}{\mbox{\sl bn}}
\newcommand{\sig}{\mbox{\sl sig}}
\newcommand{\IF}{\mbox{\mathem if}}
\newcommand{\OTHERWISE}{\mbox{\mathem otherwise}}
\newcommand{\expand}{\prec}
\newcommand{\weakexpand}{\prec^W}
\newcommand{\spcomma}{~,~}

@@

We propose DOT as a core calculus for path-dependent types. We present
the calculus formally in section \ref{dot-calculus} and through examples in
section~\ref{dot-examples}. Though we show that the calculus does not
satisfy the standard theorem of preservation in
section~\ref{dot-preservation}, we contribute a proof sketch of type
safety using logical relations in section~\ref{dot-type-safety}. In
section~\ref{discussion}, we discuss choices and variants of the
calculus, as well as related work, and conclude in
section~\ref{conclusion}.

@@

\boxfig{dot-one}{The DOT Calculus : Syntax, Reduction, Type / Declaration Assignment}{
{\bf Syntax}\medskip
\begin{center}    
$\ba{l@{\hspace{0.2mm}}|@{\hspace{0.2mm}}l}
\ba[t]{l@{\hspace{10mm}}l}
x, y, z    & \lindent{\mbox{Variable}} \\
l          & \lindent{\mbox{Value label}}\\
m          & \lindent{\mbox{Method label}}\\[0.2em]
v ::=      & \lindent{\mbox{Value}} \\
\gap x     & \mbox{variable} \\[0.2em]
t ::=      & \lindent{\mbox{Term}} \\
\gap v     & \mbox{value} \\
\gap \new x c t & \mbox{new instance} \\
\gap t.l  & \mbox{field selection} \\
\gap \mapp t m t  & \mbox{method invocation} \\[0.2em]
p ::= & \lindent \mbox{Path} \\
\gap x & \mbox{variable} \\
\gap p.l & \mbox{selection} \\
c ::= T_c \ldefs{\seq{d}} & \lindent{\mbox{Constructor}} \\[0.2em]
d ::= & \lindent{\mbox{Initialization}}\\
\gap l = v & \mbox{field initialization}\\
\gap m(x) = t & \mbox{method initialization}\\
s      ::= \seq{x \mapsto c} & \lindent\mbox{Store}
\ea
&
\ba[t]{l@{\hspace{10mm}}l}
L ::=      & \lindent{\mbox{Type label}} \\
\gap L_c   & \mbox{class label} \\
\gap L_a   & \mbox{abstract type label} \\[0.2em]
S,T,U,V,W ::= & \lindent\mbox{Type}\\
\gap p.L & \mbox{type selection} \\
\gap T \refine z {\seq D} & \mbox{refinement} \\
\gap T \tand T & \mbox{intersection type} \\
\gap T \tor T & \mbox{union type} \\
\gap \Top  & \mbox{top type} \\
\gap \Bot  & \mbox{bottom type} \\[0.2em]
S_c, T_c ::= & \lindent \mbox{Concrete type} \\
\multicolumn{2}{l}{\gap p.L_c ~|~ T_c \refine z {\seq D} ~|~ T_c \wedge T_c  ~|~ \Top} \\[0.2em]
D ::= & \lindent\mbox{Declaration} \\
\gap \Ldecl L S U & \mbox{type declaration} \\
\gap \ldecl l T   & \mbox{value declaration} \\
\gap \mdecl m S U & \mbox{method declaration}\\
\ \\
\Gamma ::= \seq{x \typ T} & \lindent\mbox{Environment}
\ea
\ea$
\end{center}
\medskip

@@

\infrule[\textsc{new}]
{
y \notin \fn(T') \\
\Gamma \ts T_c \wfe \spcomma T_c \expand_y \seq{D} \\
\Gamma \envplus{y: T_c} \ts \seq{d} \typ \seq{D} \spcomma {t'} \typ {T'}}
{\Gamma \ts \new y {T_c \ldefs{\seq{d}}} {t'} \typ T'}

@@

\begin{multicols}{2}[\judgement{Declaration Assignment}{\fbox{$\Gamma \ts d \typ D$}}]

\infrule[\textsc{vdecl}]
{\Gamma \ts v \typ {V'} \spcomma {V'} \sub V}
{\Gamma \ts l = v \typ \ldecl l V}

\infrule[\textsc{mdecl}]
{\Gamma \ts S \wfe\\
 \Gamma \envplus{x : S} \ts t \typ {T'} \spcomma {T'} \sub T}
{\Gamma \ts m(x) = t \typ \mdecl m S T }

\begin{multicols}{2}[\judgement{}{\fbox{$\Gamma \ts \seq{d} \typ \seq{D}$}}]

\infrule[\textsc{decls}]
{\Gamma \ts \seq{d_i \typ D_i} \spcomma \seq{S \sub U}}
{\Gamma \ts \seq{d} \typ \seq{D}, \seq{\Ldecl L S U}}
